\textbf{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

### 2024-01-24 manuscript_BiSC_outdoor_AP.tex
Updates from Overleaf

\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, and PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models. The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}. All models exhibit relatively good performance, evaluated by the $R^{2}_{\text{10-CV}}$, $RMSE_{\text{10-CV}}$, $R^{2}_{\text{LOOCV}}$, and $RMSE_{\text{LOOCV}}$.

The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{10-CV}} = 0.66$), and BC ($R^{2}_{\text{10-CV}} = 0.86$), with corresponding $RMSE_{\text{10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively. However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{10-CV}} = 0.44$).

In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{LOOCV}} = 0.45$), and BC ($R^{2}_{\text{LOOCV}} = 0.83$) exposure concentrations.

Dispersion models were evaluated against observations from the XVPC monitoring stations, we average the metrics obtained during this process to gauge the overall performance of the dispersion model. A comparison between predicted and observed values for all the monitoring stations used in the DM is provided in the \textbf{supplementary materials}.

The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations. LUR models indicate high variability in exposure concentrations during pregnancy, whereas DM exhibit less variability. HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.

%%% --- Figure. Average exposure during the entire pregnancy period --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure1.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\label{fig2}
\end{figure}

\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
\label{fig3}
\end{figure}

\newpage
\subsection{Comparison between models} % broad title then change this
% --- Exposure concentrations --- %
Predictions from outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig4}. The exposure range was larger for the LUR model for all the pollutants assessed. Dispersion model predicted lower exposure concentration values than LUR and Hybrid. In addition to that HM show less variability in the exposure concentration.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Correlation between measures and model estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 4.A)}. Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 4.B)}. Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure. Nitrogen dioxide

>>>>

\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models. The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}. In general, LUR and HM models exhibit relatively good performance, while DM shown a more conservative performance evaluated by the $R^{2}$ and $RMSE$.

The HM demonstrated moderate to good performance for NO\textsubscript{2} ($R^{2}_{\text{ 10-CV}} = 0.64$), PM\textsubscript{2.5} ($R^{2}_{\text{ 10-CV}} = 0.66$), and BC ($R^{2}_{\text{ 10-CV}} = 0.86$), with corresponding $RMSE_{\text{ 10-CV}}$ values of $7.48 \, \mu\text{g/m}^{3}$, $3.45 \, \mu\text{g/m}^{3}$, and $0.23 \, \mu\text{g/m}^{3}$, respectively. However, for PM\textsubscript{2.5} components, the performance was more conservative, with PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ 10-CV}} = 0.54$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ 10-CV}} = 0.70$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ 10-CV}} = 0.44$).

In contrast, LUR models showed better performance in predicting PM\textsubscript{2.5 - Fe} ($R^{2}_{\text{ LOOCV}} = 0.89$), PM\textsubscript{2.5 - Cu} ($R^{2}_{\text{ LOOCV}} = 0.87$), and PM\textsubscript{2.5 - Zn} ($R^{2}_{\text{ LOOCV}} = 0.85$) concentration levels in comparison to  NO\textsubscript{2} ($R^{2}_{\text{ LOOCV}} = 0.62$), PM\textsubscript{2.5} ($R^{2}_{\text{ LOOCV}} = 0.45$), and BC ($R^{2}_{\text{ LOOCV}} = 0.83$) exposure concentrations.

Dispersion models, when compared with observations from the XVPCA monitoring stations were evaluated against observations from the XVPCA monitoring stations, yielded an $R^{2} = 0.39$ for NO\textsubscript{2}, $R^{2} = 0.26$ for PM\textsubscript{2.5}, and $R^{2} = 0.65$ for BC. A comparison between predicted and observed values for all the monitoring stations used in the DM is provided in the \textbf{supplementary materials}.

The maps reveals distinct patterns of variability in estimated outdoor air pollution exposure concentrations. LUR models shown a high variability in exposure concentrations during pregnancy, whereas DM exhibit less variability. HM displays a variability pattern in the estimated air pollution exposures similar to those described for LUR models.

%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure1.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\label{fig2}
\end{figure}

%%% --- Figure 4. Average exposure during the entire pregnancy period --- %%%
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
\label{fig3}
\end{figure}

Results from the external validation of the LUR, DM, and HM models, conducted where information was available is depicted in \Cref{table2}. For NO\textsubscript{2}

%%% --- Table 2. Performance from external validation for LUR, DM, HM --- %%%

\begin{table}
    \centering
    \caption{Performance from external validation for LUR, DM, and HM.}
    \label{table2}
    \begin{tabular}{llllllll}
    \toprule
         Model & Validation dataset & \multicolumn{2}{c}{NO\textsubscript{2}} &  \multicolumn{2}{c}{PM\textsubscript{2.5}} &  \multicolumn{2}{c}{BC} \\
         \cmidrule(lr){3-8}
         & & R\textsuperscript{2} & RMSE & R\textsuperscript{2} & RMSE & R\textsuperscript{2} & RMSE \\
    \midrule
         LUR & External Validation & - & - & - & - & - & - \\
         DM & External Validation & - & - & - & - & - & - \\
         HM & External Validation & - & - & - & - & - & - \\
    \bottomrule
    \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Comparison between models --- %%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\subsection{Comparison between models} % broad title then change this

% --- Exposure concentrations --- %
Estimated outdoor air pollution concentrations exposure during the entire pregnancy for each pollutant and model is shown in \Cref{fig4}. The exposure range was larger for the LUR model for all the pollutants assessed. Dispersion model predicted lower exposure concentration values than LUR and Hybrid. In addition to that HM show less variability in the exposure concentration.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --- Correlation between measures and model estimates --- %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Correlation for long-term exposure (entire pregnancy) at the residential addresses ranged from r = 0.48 to 0.92 \textbf{(figure 4.A)}. Correlation for short-term exposure (weekly exposure) varied from r = 0.37 to 0.90 \textbf{(figure 4.B)}. Correlation between the air pollutant estimates for the different models assessed was moderate to strongly correlated for both long-and-short term exposure.

NO\textsubscript{2}
